---
title: "Kavita ICI CKD — Analysis"
author: "Tianqi Ouyang"
date: "`r Sys.Date()`"
execute:
  eval: true
  warning: false
  message: false
---

# 1 Setup

```{r}
#| label: setup-analysis
library(tidyverse)
library(readxl)
library(scales)
library(ggplot2)
library(gtsummary)
library(table1)
library(survival)
library(tidycmprsk)
library(ggsurvfit)

# Packages for exporting editable vector plots to PowerPoint
for (pkg in c("officer", "rvg")) {
  if (!requireNamespace(pkg, quietly = TRUE))
    install.packages(pkg, repos = "https://cloud.r-project.org")
}
library(officer)
library(rvg)

# Helper: save a plot as an editable vector slide in PowerPoint.
# Uses dml(code = print(...)) so it handles composite plots (risk tables, patchwork).
# Output lives in figures-ppt/ and is copied into _site/ by Quarto (see _quarto.yml resources).
save_plot_pptx <- function(plot_obj, name, width = 10, height = 6) {
  dir.create("figures-ppt", showWarnings = FALSE, recursive = TRUE)
  path <- file.path("figures-ppt", paste0(name, ".pptx"))
  # Force the plot into a local variable so the code block is self-contained
  local_plot <- plot_obj
  doc <- read_pptx()
  doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")
  doc <- ph_with(doc,
                 dml(code = print(local_plot)),
                 location = ph_location(left = 0.25, top = 0.25,
                                        width = width, height = height))
  print(doc, target = path)
  invisible(path)
}

# Cleaned master produced by the data management pipeline (see data_management.qmd)
master_updated <- read_excel(
  "Final cohort/final_master_0114_raw.xlsx") %>%
  mutate(across(contains("date"), as.Date))

aki <- read_excel("Final cohort/aki_0120.xlsx") %>%
  mutate(across(contains("date"), as.Date))

# ICI class reference
ici_class <- read_excel("ICI CKD/final_master_1229.xlsx") %>%
  select(EMPI, ICI_Class)
```

# 1b Analytical derivations

Everything below duplicates the derivation block at the top of the original
`Kavita cohort 2026 0130.Rmd`. It takes the cleaned master + AKI table and
builds the composite outcomes, competing-risk status, yearly eGFR / CKD stage,
and scaled regression covariates used in every downstream section.

```{r}
#| label: fn-ckd-epi-gfr
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  gfr <- 142 * min(c(creat/k, 1))^a *
               max(c(creat/k, 1))^-1.2 *
               0.9938^age * m
  return(gfr)
}
gfr_calc <- Vectorize(ckd_epi_gfr)

get_ckd_stage <- function(egfr) {
  if (is.na(egfr))     return(NA)
  else if (egfr >= 90) return(1)
  else if (egfr >= 60) return(2)
  else if (egfr >= 30) return(3)
  else if (egfr >= 15) return(4)
  else                 return(5)
}
```

```{r}
#| label: merge-aki-ici
final_master <- master_updated %>%
  merge(aki, by = "EMPI", all.x = TRUE) %>%
  mutate(aki_days = as.numeric(aki_date - ICI_Dose_Date),
         aki      = if_else(is.na(aki), 0, aki)) %>%
  merge(ici_class, by = "EMPI", all.x = TRUE) %>%
  filter(is.na(Date_Of_Death) | Date_Of_Death >= ICI_Dose_Date)
```

```{r}
#| label: cancer-groups-race
final_master <- final_master %>%
  mutate(cancer_group = case_when(
    cancer_type == "gi"                   ~ "gi",
    cancer_type == "cutaneous / melanoma" ~ "cutaneous / melanoma",
    cancer_type == "gu"                   ~ "gu",
    cancer_type == "gyn"                  ~ "gyn",
    cancer_type == "thoracic"             ~ "thoracic",
    TRUE                                   ~ "other"
  )) %>%
  mutate(
    Race1 = ifelse(is.na(Race1), "Other/Unknown", Race1),
    Race  = case_when(
      Race1 == "White" ~ "White",
      Race1 == "Black" ~ "Black",
      Race1 == "Asian" ~ "Asian",
      TRUE             ~ "Other/Unknown"
    ),
    Ethnic_Group = case_when(
      Ethnic_Group == "HISPANIC"     ~ "Hispanic",
      Ethnic_Group == "Non Hispanic" ~ "Non_hispanic",
      TRUE                           ~ "Other"
    ),
    Ethnic_Group = case_when(
      Race != "Black" & Language == "Spanish" ~ "Hispanic",
      TRUE                                     ~ Ethnic_Group
    )
  )
```

```{r}
#| label: composite-outcomes
final_master <- final_master %>%
  mutate(
    eskd_composite      = ifelse(esrd_kt_after_ici == 1 | eskd_defination_1 == 1, 1, 0),
    eskd_composite      = ifelse(is.na(eskd_composite), 0, eskd_composite),
    eskd_composite_date = pmin(esrd_kt_after_ici_date, eskd_defination_1_date, na.rm = TRUE),
    time_to_eskd_composite  = as.numeric(eskd_composite_date  - ICI_Dose_Date),
    time_to_ckd_incidence   = as.numeric(ckd_incidence_date   - ICI_Dose_Date),
    time_to_ckd_progression = as.numeric(ckd_progression_date - ICI_Dose_Date),

    ckd_composite = ifelse(eskd_composite == 1 | ckd_incidence == 1 |
                           ckd_progression == 1, 1, 0),
    ckd_composite = ifelse(is.na(ckd_composite), 0, ckd_composite),
    time_to_ckd_composite = pmin(time_to_eskd_composite,
                                 time_to_ckd_incidence,
                                 time_to_ckd_progression, na.rm = TRUE),

    last_cre_days        = as.numeric(last_cre_date - ICI_Dose_Date),
    death_days           = as.numeric(Date_Of_Death - ICI_Dose_Date),
    time_to_death        = death_days,
    time_to_death_status = ifelse(!is.na(death_days), 2, 0),

    status_ckd_cmp = case_when(
      (ckd_composite == 1) & (time_to_death_status == 2) ~
        ifelse(time_to_ckd_composite < time_to_death, 1, 2),
      (ckd_composite == 1 & time_to_death_status == 0) ~ 1,
      (ckd_composite == 0 & time_to_death_status == 2) ~ 2,
      TRUE                                              ~ 0
    ),
    time_to_ckd_cmp = case_when(
      status_ckd_cmp == 1 ~ time_to_ckd_composite,
      status_ckd_cmp == 2 ~ time_to_death,
      status_ckd_cmp == 0 ~ last_cre_days
    ),
    status_ckd_cmp = factor(status_ckd_cmp, levels = c(0, 1, 2),
                            labels = c("Censored", "CKD Composite", "Death")),
    time_to_ckd_cmp_year = time_to_ckd_cmp / 365,
    status_ckd_csh       = ifelse(status_ckd_cmp == "CKD Composite", 1, 0),
    aki_before_ckd       = if_else(aki_days <= time_to_ckd_composite, 1, 0)
  )
```

```{r}
#| label: ici-class
final_master <- final_master %>%
  mutate(ICI_Class = case_when(
    !is.na(ICI_Class) ~ ICI_Class,
    ICI_Name %in% c("pembrolizumab","nivolumab","cemiplimab",
                    "Nivolumab","Dostarlimab","Cemiplimab") ~ "PD1",
    ICI_Name %in% c("atezolizumab","avelumab","durvalumab")  ~ "PDL1",
    ICI_Name == "ipilimumab"                                  ~ "CTLA4",
    ICI_Name == "ipilimumab AND nivolumab"                    ~ "Combination",
    TRUE                                                       ~ NA_character_
  ))
```

```{r}
#| label: yearly-egfr
followed_cols <- setNames(
  lapply(1:10, function(y) ifelse(final_master$last_cre_days > 365 * y, 1, 0)),
  paste0("followed_", 1:10, "year"))
final_master <- bind_cols(final_master, as.data.frame(followed_cols))

for (y in 1:10) {
  col_cre  <- paste0("pre_mdeian_CRE_", y, "year")
  col_egfr <- paste0("eGFR_CRE_",       y, "year")
  if (col_cre %in% names(final_master)) {
    final_master[[col_egfr]] <- mapply(gfr_calc,
                                       final_master[[col_cre]],
                                       final_master$male,
                                       final_master$age_ici)
  }
}

final_master <- final_master %>%
  mutate(across(any_of(paste0("eGFR_CRE_", 1:10, "year")),
                ~ coalesce(., eGFR_CRE_baseline)))

final_master$ckd_stage_baseline <- sapply(final_master$eGFR_CRE_baseline, get_ckd_stage)
for (y in 1:10) {
  col_egfr  <- paste0("eGFR_CRE_",  y, "year")
  col_stage <- paste0("ckd_stage_", y, "year")
  if (col_egfr %in% names(final_master)) {
    final_master[[col_stage]] <- sapply(final_master[[col_egfr]], get_ckd_stage)
  }
}
```

```{r}
#| label: scaling-and-year-flags
final_master <- final_master %>%
  mutate(
    year_10                       = age_ici / 10,
    eGFR_CRE_baseline_decrease_10 = (eGFR_CRE_baseline * -1) / 10,
    nephrotoxic_chemo = if_else(
      if_any(c(Bevacizumab, Cisplatin, Carboplatin, Pemetrexed,
               Gemcitabine, Cetuximab, Trastuzumab), ~ . == 1),
      1L, 0L)
  )

for (y in 1:10) {
  col_egfr <- paste0("eGFR_CRE_", y, "year")
  col_dec  <- paste0("eGFR_CRE_", y, "year_decrease_10")
  if (col_egfr %in% names(final_master)) {
    final_master[[col_dec]] <- (final_master[[col_egfr]] * -1) / 10
  }
}

for (y in 1:10) {
  d <- 365 * y
  final_master[[paste0("eskd_composite_",  y, "year")]] <- ifelse(final_master$time_to_eskd_composite   < d, 1, 0)
  final_master[[paste0("ckd_incidence_",   y, "year")]] <- ifelse(final_master$time_to_ckd_incidence    < d, 1, 0)
  final_master[[paste0("ckd_progression_", y, "year")]] <- ifelse(final_master$time_to_ckd_progression  < d, 1, 0)
  final_master[[paste0("ckd_composite_",   y, "year")]] <- ifelse(final_master$time_to_ckd_composite    < d, 1, 0)
}

final_master <- final_master %>%
  mutate(across(matches("(ckd_incidence|ckd_progression|eskd_composite|ckd_composite).*year$"),
                ~ replace_na(., 0))) %>%
  mutate(across(matches("(ckd_incidence|ckd_progression|eskd_composite).*year$"),
                ~ factor(replace_na(., 0))))
```

```{r}
#| label: factor-conversion
num_var <- c("age_ici","pre_CRE_180days","year_10",
             "pre_HGB_180days","pre_ALB_180days","eGFR_CRE_baseline")

cat_var_candidates <- c("Gender_Legal_Sex","male","Race","Ethnic_Group",
  "ckd_incidence","ckd_progression","eskd_composite","ckd_composite",
  paste0("ckd_stage_", 1:10, "year"),
  "dm","htn","cad","cirrhosis","esrd_kt",
  "ace_arb","diu","ppi","steroids","smoking","statins",
  "ICI_Class","nephrotoxic_chemo","ckd_stage_baseline",
  paste0("ckd_composite_", 1:5, "year"),
  "Bevacizumab","Cisplatin","Carboplatin","Pemetrexed",
  "Gemcitabine","Cetuximab","Trastuzumab","VEGF_TKI",
  "cancer_type","cancer_group","aki")

final_master <- final_master %>%
  mutate_at(intersect(cat_var_candidates, names(.)), factor)

if ("ICI_Class" %in% names(final_master) && is.factor(final_master$ICI_Class)) {
  final_master$ICI_Class <- relevel(final_master$ICI_Class, ref = "PD1")
}
```

# 2 Variable Lists & Helpers

```{r}
#| label: variable-lists
all_var <- c(
  # Demographics
  "age_ici", "year_10", "Gender_Legal_Sex", "male", "Race", "Ethnic_Group",

  # Outcomes
  "ckd_incidence", "ckd_progression", "eskd_composite", "ckd_composite",

  # Longitudinal CKD stage
  paste0("ckd_stage_", 1:10, "year"),

  # Comorbidities
  "dm", "htn", "cad", "cirrhosis", "esrd_kt",

  # Medications
  "ace_arb", "diu", "ppi", "steroids", "smoking", "statins",
  "ICI_Class", "nephrotoxic_chemo",

  # Labs / kidney function
  "pre_CRE_180days", "pre_HGB_180days", "pre_ALB_180days",
  "eGFR_CRE_baseline", "ckd_stage_baseline",

  # Composite outcome flags
  paste0("ckd_composite_", 1:5, "year"),

  # Individual chemo agents
  "Bevacizumab", "Cisplatin", "Carboplatin", "Pemetrexed",
  "Gemcitabine", "Cetuximab", "Trastuzumab", "VEGF_TKI",

  # Cancer
  "cancer_type", "cancer_group",

  # AKI
  "aki"
)

num_var <- c("age_ici", "year_10",
             "pre_CRE_180days", "pre_HGB_180days", "pre_ALB_180days",
             "eGFR_CRE_baseline")

cat_var <- setdiff(all_var, num_var)
```

```{r}
#| label: fn-create-outcome-table
# Build a year-specific tbl_summary with cumulative CKD stage history
create_outcome_table <- function(data, time_suffix) {
  year_num           <- as.numeric(gsub("[^0-9]", "", time_suffix))
  stage_history_cols <- paste0("ckd_stage_", 1:year_num, "year")

  followed_col  <- paste0("followed_",       time_suffix)
  ckd_inc_col   <- paste0("ckd_incidence_",  time_suffix)
  ckd_prog_col  <- paste0("ckd_progression_", time_suffix)
  eskd_comp_col <- paste0("eskd_composite_", time_suffix)
  ckd_comp_col  <- paste0("ckd_composite_",  time_suffix)

  data %>%
    filter(.data[[followed_col]] == 1) %>%
    select(
      # Demographics
      age_ici, male, Race, Ethnic_Group,

      # Outcomes (dynamic)
      all_of(ckd_inc_col), all_of(ckd_prog_col),
      all_of(eskd_comp_col), all_of(ckd_comp_col),

      # Longitudinal CKD stage (cumulative up to the year of interest)
      all_of(stage_history_cols), ckd_stage_baseline,
      aki,

      # Comorbidities
      dm, htn, cad, cirrhosis,

      # Medications
      ace_arb, diu, ppi, steroids, smoking, statins,
      ICI_Class, nephrotoxic_chemo,
      cancer_type, cancer_group,

      # Labs
      pre_CRE_180days, pre_HGB_180days, pre_ALB_180days,
      eGFR_CRE_baseline
    ) %>%
    tbl_summary(
      missing_text = "Missing",
      statistic    = list(all_continuous() ~ "{mean} ({sd})")
    )
}
```

```{r}
#| label: year-cohorts
final_master_1year <- final_master %>% filter(followed_1year == 1)
final_master_2year <- final_master %>% filter(followed_2year == 1)
final_master_3year <- final_master %>% filter(followed_3year == 1)
final_master_4year <- final_master %>% filter(followed_4year == 1)
final_master_5year <- final_master %>% filter(followed_5year == 1)
```

---

# 3 Baseline Characteristics

## 3.0 Distribution check for continuous variables

Before choosing summary statistics for Table 1, assess whether `age_ici`,
`pre_CRE_180days` (baseline creatinine), and `eGFR_CRE_baseline` are
approximately normal. With N ≈ 17,700 the Shapiro-Wilk test is not applicable
(N > 5000), so we rely on **visual distribution + skewness / kurtosis**.

```{r}
#| label: normality-setup
#| include: false
if (!requireNamespace("moments",  quietly = TRUE)) install.packages("moments", repos = "https://cloud.r-project.org")
if (!requireNamespace("patchwork", quietly = TRUE)) install.packages("patchwork", repos = "https://cloud.r-project.org")
library(moments)
library(patchwork)
```

```{r}
#| label: normality-stats
check_vars <- c("age_ici", "pre_CRE_180days", "eGFR_CRE_baseline")
diag_df <- final_master %>%
  select(all_of(check_vars)) %>%
  mutate(across(everything(), ~ suppressWarnings(as.numeric(as.character(.)))))

dist_stats <- purrr::map_dfr(check_vars, function(v) {
  x <- diag_df[[v]]; x <- x[is.finite(x)]
  data.frame(
    Variable         = v,
    N                = length(x),
    Mean             = round(mean(x), 2),
    SD               = round(sd(x), 2),
    Median           = round(median(x), 2),
    `Q1 - Q3`        = paste0(round(quantile(x, 0.25), 2), " – ",
                              round(quantile(x, 0.75), 2)),
    Skewness         = round(moments::skewness(x), 2),
    `Kurtosis (excess)` = round(moments::kurtosis(x) - 3, 2),
    check.names = FALSE
  )
})
knitr::kable(dist_stats,
             caption = "Distribution statistics. |Skewness| < 0.5 → approximately normal; 0.5–1 → mild skew; > 1 → strongly skewed.")
```

```{r}
#| label: normality-plots
#| fig-width: 11
#| fig-height: 9
make_plots <- function(x, lab) {
  x <- x[is.finite(x)]
  d <- data.frame(x = x)
  p1 <- ggplot(d, aes(x)) +
    geom_histogram(bins = 40, fill = "#4C72B0", color = "white") +
    labs(title = paste0("Histogram — ", lab), x = lab, y = "Count") +
    theme_classic()
  p2 <- ggplot(d, aes(sample = x)) +
    stat_qq(color = "#4C72B0", alpha = 0.4, size = 0.5) +
    stat_qq_line(color = "red", linewidth = 0.6) +
    labs(title = paste0("Q-Q plot — ", lab),
         x = "Theoretical quantiles", y = "Sample quantiles") +
    theme_classic()
  p1 | p2
}

(make_plots(diag_df$age_ici,           "Age at ICI start (years)")) /
(make_plots(diag_df$pre_CRE_180days,   "Baseline creatinine (mg/dL)")) /
(make_plots(diag_df$eGFR_CRE_baseline, "Baseline eGFR (mL/min/1.73m²)"))
```

```{r}
#| label: normality-decision
# Rule: |skewness| ≤ 0.5 → treat as approximately normal → mean (SD) in Table 1
normal_flag    <- setNames(abs(dist_stats$Skewness) <= 0.5, dist_stats$Variable)
normal_vars    <- names(normal_flag)[ normal_flag]
nonnormal_vars <- names(normal_flag)[!normal_flag]
```

```{r}
#| label: normality-decision-show
#| echo: false
#| results: asis
cat("**Decision:**\n\n",
    "- Reported as **mean (SD)** in Table 1: ",
    if (length(normal_vars))    paste0("`", normal_vars,    "`", collapse = ", ") else "_none_",
    "\n- Reported as **median (Q1, Q3)**: ",
    if (length(nonnormal_vars)) paste0("`", nonnormal_vars, "`", collapse = ", ") else "_none_",
    "\n", sep = "")
```

## 3.1 Overall Cohort

```{r}
#| label: table1-overall
# All continuous variables reported as mean (SD)
final_master %>%
  select(
    age_ici, male, Race, Ethnic_Group,
    ckd_progression, ckd_incidence, eskd_composite, ckd_composite,
    ckd_stage_baseline, aki,
    dm, htn, cad, cirrhosis, cancer_type, cancer_group,
    ace_arb, diu, ppi, steroids, smoking,
    Bevacizumab, Cisplatin, Carboplatin, Pemetrexed,
    Gemcitabine, Cetuximab, Trastuzumab, VEGF_TKI,
    statins, ICI_Class, nephrotoxic_chemo,
    pre_CRE_180days, pre_HGB_180days, pre_ALB_180days, eGFR_CRE_baseline
  ) %>%
  tbl_summary(
    missing_text = "Missing",
    statistic    = list(all_continuous() ~ "{mean} ({sd})")
  )
```

## 3.2 By Year of Follow-up

```{r}
#| label: table1-1year
create_outcome_table(final_master, "1year")
```

```{r}
#| label: table1-2year
create_outcome_table(final_master, "2year")
```

```{r}
#| label: table1-3year
create_outcome_table(final_master, "3year")
```

```{r}
#| label: table1-4year
create_outcome_table(final_master, "4year")
```

```{r}
#| label: table1-5year
create_outcome_table(final_master, "5year")
```

```{r}
#| label: table1-6year
create_outcome_table(final_master, "6year")
```

```{r}
#| label: table1-7year
create_outcome_table(final_master, "7year")
```

```{r}
#| label: table1-8year
create_outcome_table(final_master, "8year")
```

```{r}
#| label: table1-9year
create_outcome_table(final_master, "9year")
```

```{r}
#| label: table1-10year
create_outcome_table(final_master, "10year")
```

---

# 4 Cumulative Incidence (CIF) — Overall

CKD Composite and Death treated as competing events.

```{r}
#| label: cif-overall
#| fig-width: 10
#| fig-height: 6
# Curves-only version — saved to PPT as editable vector
p_cif_overall_curves <- cuminc(Surv(time_to_ckd_cmp, status_ckd_cmp) ~ 1,
                               data = final_master) %>%
  ggcuminc(outcome = c("CKD Composite", "Death")) +
  add_confidence_interval() +
  labs(x = "Years", y = "Cumulative Incidence") +
  scale_ggsurvfit() +
  scale_x_continuous(
    breaks = seq(0, 365 * 5, by = 365),
    labels = 0:5,
    limits = c(0, 365 * 5)
  )

# Display version — includes risk table
p_cif_overall_curves + add_risktable()

# Save editable PPT (curves only — risk tables are composite layouts
# that can't be converted to editable vector)
save_plot_pptx(p_cif_overall_curves, "cif-overall", width = 10, height = 6)
```

[📥 Download as editable PowerPoint](figures-ppt/cif-overall.pptx){.btn .btn-primary .btn-sm download="cif-overall.pptx"}

---

# 5 CKD Stage vs Yearly CKD Composite

## 5.1 Year 1

```{r}
#| label: stage-vs-outcome-1year
table1(~ ckd_stage_baseline + factor(ckd_stage_1year) | ckd_composite_1year,
       final_master_1year)
```

## 5.2 Year 2

```{r}
#| label: stage-vs-outcome-2year
table1(~ ckd_stage_baseline + factor(ckd_stage_2year) | ckd_composite_2year,
       final_master_2year)
```

## 5.3 Year 3

```{r}
#| label: stage-vs-outcome-3year
table1(~ ckd_stage_baseline + factor(ckd_stage_3year) | ckd_composite_3year,
       final_master_3year)
```

## 5.4 Year 4

```{r}
#| label: stage-vs-outcome-4year
table1(~ ckd_stage_baseline + factor(ckd_stage_4year) | ckd_composite_4year,
       final_master_4year)
```

## 5.5 Year 5

```{r}
#| label: stage-vs-outcome-5year
table1(~ ckd_stage_baseline + factor(ckd_stage_5year) | ckd_composite_5year,
       final_master_5year)
```

---

# 6 AKI Analyses

## 6.1 AKI in 5-year CKD Composite cases

```{r}
#| label: aki-5year-ckd-cases
table1(~ aki + as.factor(aki_before_ckd),
       final_master_5year %>% filter(ckd_composite_5year == 1))
```

## 6.2 AKI overall

```{r}
#| label: aki-overall
table1(~ aki + as.factor(aki_before_ckd) + ckd_composite, final_master)
```

## 6.3 CKD Composite cases — stratified by AKI timing

```{r}
#| label: aki-before-ckd-table
final_master_ckd <- final_master %>% filter(ckd_composite == 1) %>%
  mutate(aki_before_ckd = if_else(is.na(aki_before_ckd), 0, aki_before_ckd))

create_outcome_table(final_master_ckd, "1year")

final_master_ckd %>%
  select(
    aki_before_ckd,
    age_ici, male, Race, Ethnic_Group,
    ckd_composite, ckd_incidence, ckd_progression,
    ckd_stage_baseline, aki,
    dm, htn, cad, cirrhosis, cancer_type, cancer_group,
    ace_arb, diu, ppi, steroids, smoking, statins,
    ICI_Class, nephrotoxic_chemo,
    pre_CRE_180days, pre_HGB_180days, pre_ALB_180days, eGFR_CRE_baseline
  ) %>%
  tbl_summary(
    missing_text = "Missing",
    by           = aki_before_ckd,
    statistic    = list(all_continuous() ~ "{mean} ({sd})")
  )
```

---

# 7 Logistic Regression — Outcome: `ckd_composite_5year`

## 7.1 Univariate

```{r}
#| label: lr-univariate
final_master_5year %>%
  select(
    year_10, male, Race,
    dm, htn, cad, cirrhosis,
    ace_arb, diu, ppi, steroids, smoking, statins,
    nephrotoxic_chemo,
    Bevacizumab, Cisplatin, Carboplatin, Pemetrexed,
    Gemcitabine, Cetuximab, Trastuzumab,
    eGFR_CRE_baseline_decrease_10,
    ICI_Class, cancer_group,
    ckd_composite_5year
  ) %>%
  tbl_uvregression(
    method        = glm,
    method.args   = list(family = binomial),
    y             = ckd_composite_5year,
    pvalue_fun    = ~ style_pvalue(.x, digits = 3),
    exponentiate  = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## 7.2 Multivariate — Model 1 (nephrotoxic_chemo composite)

```{r}
#| label: lr-multivariate-1
glm(
  ckd_composite_5year ~ year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo + ICI_Class + cancer_group,
  data = final_master_5year, family = binomial()
) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun   = ~ style_pvalue(.x, digits = 3)) %>%
  bold_p() %>%
  bold_labels() %>%
  add_n()
```

## 7.3 Multivariate — Model 2 (individual chemo agents)

```{r}
#| label: lr-multivariate-2
glm(
  ckd_composite_5year ~ year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 +
    Bevacizumab + Cisplatin + Carboplatin + Pemetrexed +
    Gemcitabine + Cetuximab + Trastuzumab +
    ICI_Class + cancer_group,
  data = final_master_5year, family = binomial()
) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun   = ~ style_pvalue(.x, digits = 3)) %>%
  bold_p() %>%
  bold_labels() %>%
  add_n()
```

## 7.4 Multivariate — Model 3 (with statins)

```{r}
#| label: lr-multivariate-3
glm(
  ckd_composite_5year ~ year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo + statins +
    ICI_Class + cancer_group,
  data = final_master_5year, family = binomial()
) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun   = ~ style_pvalue(.x, digits = 3)) %>%
  bold_p() %>%
  bold_labels() %>%
  add_n()
```

---

# 8 Fine-Gray Competing Risk Model (CKD Composite vs Death)

## 8.1 Univariate

```{r}
#| label: fg-univariate
final_master %>%
  select(
    year_10, male, Race,
    dm, htn, cad, cirrhosis,
    ace_arb, diu, ppi, steroids, smoking, statins,
    nephrotoxic_chemo,
    Bevacizumab, Cisplatin, Carboplatin, Pemetrexed,
    Gemcitabine, Cetuximab, Trastuzumab,
    eGFR_CRE_baseline_decrease_10,
    ICI_Class, cancer_group,
    time_to_ckd_cmp, status_ckd_cmp
  ) %>%
  tbl_uvregression(
    method       = tidycmprsk::crr,
    method.args  = list(failcode = "CKD Composite", cencode = "Censored"),
    y            = Surv(time_to_ckd_cmp, status_ckd_cmp),
    pvalue_fun   = ~ style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## 8.2 Multivariate — Model 1 (individual chemo agents)

```{r}
#| label: fg-multivariate-1
tidycmprsk::crr(
  Surv(time_to_ckd_cmp, status_ckd_cmp) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 +
    Bevacizumab + Cisplatin + Carboplatin + Pemetrexed +
    Gemcitabine + Cetuximab + Trastuzumab +
    ICI_Class + cancer_group + cirrhosis,
  failcode = "CKD Composite", cencode = "Censored",
  data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## 8.3 Multivariate — Model 2 (nephrotoxic_chemo composite)

```{r}
#| label: fg-multivariate-2
tidycmprsk::crr(
  Surv(time_to_ckd_cmp, status_ckd_cmp) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo +
    ICI_Class + cancer_group + cirrhosis,
  failcode = "CKD Composite", cencode = "Censored",
  data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## 8.4 Multivariate — Model 3 (with statins)

```{r}
#| label: fg-multivariate-3
tidycmprsk::crr(
  Surv(time_to_ckd_cmp, status_ckd_cmp) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo + statins +
    ICI_Class + cancer_group + cirrhosis,
  failcode = "CKD Composite", cencode = "Censored",
  data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

---

# 9 Cause-Specific Cox Model (CKD Composite as the event of interest)

## 9.1 Univariate

```{r}
#| label: csh-univariate
final_master %>%
  select(
    year_10, male, Race,
    dm, htn, cad, cirrhosis,
    ace_arb, diu, ppi, steroids, smoking, statins,
    nephrotoxic_chemo,
    eGFR_CRE_baseline_decrease_10,
    ICI_Class, cancer_group,
    time_to_ckd_cmp, status_ckd_csh
  ) %>%
  tbl_uvregression(
    method       = survival::coxph,
    y            = Surv(time_to_ckd_cmp, status_ckd_csh),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## 9.2 Multivariate — Model 1 (individual chemo agents)

```{r}
#| label: csh-multivariate-1
survival::coxph(
  Surv(time_to_ckd_cmp, status_ckd_csh) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 +
    Bevacizumab + Cisplatin + Carboplatin + Pemetrexed +
    Gemcitabine + Cetuximab + Trastuzumab +
    ICI_Class + cancer_group + cirrhosis,
  data = final_master
) %>%
  tbl_regression(exponentiate = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## 9.3 Multivariate — Model 2 (nephrotoxic_chemo composite)

```{r}
#| label: csh-multivariate-2
survival::coxph(
  Surv(time_to_ckd_cmp, status_ckd_csh) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo +
    ICI_Class + cancer_group + cirrhosis,
  data = final_master
) %>%
  tbl_regression(exponentiate = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## 9.4 Multivariate — Model 3 (with statins)

```{r}
#| label: csh-multivariate-3
survival::coxph(
  Surv(time_to_ckd_cmp, status_ckd_csh) ~
    year_10 + male + dm + htn + cad + ppi + smoking +
    eGFR_CRE_baseline_decrease_10 + nephrotoxic_chemo + statins +
    ICI_Class + cancer_group + cirrhosis,
  data = final_master
) %>%
  tbl_regression(exponentiate = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

---

# 10 CIF Plots — Stratified

## 10.1 By Diabetes Status

```{r}
#| label: cif-dm
#| fig-width: 11
#| fig-height: 7
final_master_dm <- final_master %>%
  mutate(dm_label = factor(dm, levels = c(0, 1),
                           labels = c("No Diabetes", "Diabetes")))

p_cif_dm_curves <- cuminc(Surv(time_to_ckd_cmp_year, status_ckd_cmp) ~ dm_label,
                          data = final_master_dm) %>%
  ggcuminc(outcome = "CKD Composite") +
  add_confidence_interval() +
  scale_color_manual(values = c("Diabetes" = "#619CFF", "No Diabetes" = "#F8766D")) +
  scale_fill_manual(values  = c("Diabetes" = "#619CFF", "No Diabetes" = "#F8766D")) +
  coord_cartesian(xlim = c(0, 5)) +
  scale_x_continuous(
    breaks = seq(0, 5, by = 1),
    limits = c(0, 5),
    expand = c(0.02, 0)
  ) +
  labs(x = "Years since Baseline",
       y = "Cumulative Incidence",
       title = "CIF of CKD Composite by DM Status (5-year view)") +
  theme_classic() +
  theme(
    legend.position   = "bottom",
    axis.text.x       = element_text(size = 8),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank()
  )

# Display with risk table
p_cif_dm_curves +
  add_risktable(
    times           = seq(0, 5, by = 1),
    risktable_stats = c("n.risk", "cum.event"),
    stats_label     = list(n.risk = "At Risk", cum.event = "Events")
  )

save_plot_pptx(p_cif_dm_curves, "cif-by-dm", width = 11, height = 7)
```

[📥 Download as editable PowerPoint](figures-ppt/cif-by-dm.pptx){.btn .btn-primary .btn-sm download="cif-by-dm.pptx"}

## 10.2 By Cancer Type

```{r}
#| label: cif-cancer
#| fig-width: 12
#| fig-height: 8
# Shorten the long "cutaneous / melanoma" label so the risk table fits cleanly
final_master_cancer <- final_master %>%
  mutate(cancer_group = fct_recode(
    factor(cancer_group),
    "Melanoma" = "cutaneous / melanoma",
    "GI"       = "gi",
    "GU"       = "gu",
    "GYN"      = "gyn",
    "Thoracic" = "thoracic",
    "Other"    = "other"
  ))

p_cif_cancer_curves <- cuminc(Surv(time_to_ckd_cmp_year, status_ckd_cmp) ~ cancer_group,
                              data = final_master_cancer) %>%
  ggcuminc(outcome = "CKD Composite") +
  add_confidence_interval() +
  coord_cartesian(xlim = c(0, 5)) +
  scale_x_continuous(
    breaks = seq(0, 5, by = 1),
    limits = c(0, 5),
    expand = c(0.02, 0)
  ) +
  labs(x = "Years since Baseline",
       y = "Cumulative Incidence",
       title = "CIF of CKD Composite by Cancer Type (5-year view)") +
  theme_classic() +
  theme(
    legend.position   = "bottom",
    axis.text.x       = element_text(size = 8),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank()
  )

# Display with risk table
p_cif_cancer_curves +
  add_risktable(
    times           = seq(0, 5, by = 1),
    risktable_stats = c("n.risk", "cum.event"),
    stats_label     = list(n.risk = "At Risk", cum.event = "Events")
  )

save_plot_pptx(p_cif_cancer_curves, "cif-by-cancer", width = 12, height = 8)
```

[📥 Download as editable PowerPoint](figures-ppt/cif-by-cancer.pptx){.btn .btn-primary .btn-sm download="cif-by-cancer.pptx"}

---

# 11 Overall Survival by Cancer Type (5-year Kaplan-Meier)

Constructs a standard KM survival curve for all-cause mortality, stratified by
`cancer_group`. Patients without a `Date_Of_Death` are censored at
`last_cre_date` (last known alive). Log-rank test assesses whether the survival
distributions differ across cancer groups.

```{r}
#| label: os-prep
final_master_os <- final_master %>%
  mutate(
    death_event      = if_else(time_to_death_status == 2, 1, 0),
    death_time_days  = if_else(death_event == 1, time_to_death, last_cre_days),
    death_time_year  = death_time_days / 365,
    cancer_group     = fct_recode(
      factor(cancer_group),
      "Melanoma" = "cutaneous / melanoma",
      "GI"       = "gi",
      "GU"       = "gu",
      "GYN"      = "gyn",
      "Thoracic" = "thoracic",
      "Other"    = "other"
    )
  ) %>%
  filter(!is.na(death_time_days), death_time_days >= 0, !is.na(cancer_group))
```

## 11.1 KM survival curves by cancer type

```{r}
#| label: os-km-plot
#| fig-width: 12
#| fig-height: 8
p_os_cancer_curves <- survfit2(Surv(death_time_year, death_event) ~ cancer_group,
                               data = final_master_os) %>%
  ggsurvfit() +
  add_confidence_interval() +
  add_pvalue(location = "annotation", x = 0.2, y = 0.10) +
  coord_cartesian(xlim = c(0, 5)) +
  scale_x_continuous(
    breaks = seq(0, 5, by = 1),
    limits = c(0, 5),
    expand = c(0.02, 0)
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1)
  ) +
  labs(
    x     = "Years since ICI Start",
    y     = "Overall Survival",
    title = "Overall Survival by Cancer Type (5-year view)"
  ) +
  theme_classic() +
  theme(
    legend.position  = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# Display with risk table
p_os_cancer_curves +
  add_risktable(
    times           = seq(0, 5, by = 1),
    risktable_stats = c("n.risk", "cum.event"),
    stats_label     = list(n.risk = "At Risk", cum.event = "Deaths")
  )

save_plot_pptx(p_os_cancer_curves, "os-by-cancer", width = 12, height = 8)
```

[📥 Download as editable PowerPoint](figures-ppt/os-by-cancer.pptx){.btn .btn-primary .btn-sm download="os-by-cancer.pptx"}

## 11.2 Log-rank test

```{r}
#| label: os-logrank
survdiff(Surv(death_time_year, death_event) ~ cancer_group,
         data = final_master_os)
```

## 11.3 Median survival by cancer type

```{r}
#| label: os-median
survfit(Surv(death_time_year, death_event) ~ cancer_group,
        data = final_master_os)
```

## 11.4 Cox model: cancer type effect on mortality (unadjusted)

```{r}
#| label: os-cox-unadjusted
coxph(Surv(death_time_year, death_event) ~ cancer_group,
      data = final_master_os) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun   = ~ style_pvalue(.x, digits = 3)) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## 11.5 Cox model: cancer type effect adjusted for core covariates

```{r}
#| label: os-cox-adjusted
coxph(
  Surv(death_time_year, death_event) ~
    cancer_group + year_10 + male + dm + htn + cad + cirrhosis +
    eGFR_CRE_baseline_decrease_10 + ICI_Class + nephrotoxic_chemo,
  data = final_master_os
) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun   = ~ style_pvalue(.x, digits = 3)) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```
