---
title: "Kavita ICI CKD — Data Management"
subtitle: "Code reference — not executed at render"
author: "Tianqi Ouyang"
date: "`r Sys.Date()`"
execute:
  eval: false
  warning: false
  message: false
---

::: {.callout-note}
## About this page
This document shows the data management pipeline as **code reference**.
Chunks are not executed at render time — they load RPDR extracts and patient-level
data which stay local. The cleaned output (`final_master_0114_raw.xlsx` +
`aki_0120.xlsx`) is produced locally and feeds [`analysis.qmd`](analysis.qmd).
:::

# 1 Variable Name Dictionary

Variables with the same name as the original ICI starts file / RPDR extract are
not listed. Only **renamed** and **newly created** variables are shown.

## 1.1 Renamed Variables

| Original | In Code |
|---|---|
| `EMPI_MRN` | `EMPI` |
| `age` (DOB-derived) | `age_ici` |
| `Tumor_Category` (raw) | `cancer_type` (cleaned from `final_master_1202.xlsx`) |

## 1.2 Demographic / Derived Identifiers

| In Code | Description |
|---|---|
| `male` | 1 if `Gender_Legal_Sex == "Male"`, else 0 |
| `age_ici` | Age at ICI start (days between `Date_of_Birth` and `ICI_Dose_Date`, /365) |
| `year_10` | `age_ici / 10` (per-10-year scaling for regression) |
| `Race` | Collapsed from `Race1`: White / Black / Asian / Other/Unknown |
| `Ethnic_Group` | Hispanic / Non_hispanic / Other (Spanish language overrides to Hispanic if not Black) |

## 1.3 Derived — eGFR & Kidney Function

| In Code | Description |
|---|---|
| `ckd_epi_gfr` | CKD-EPI 2021 race-free creatinine-based eGFR function |
| `eGFR_cre` | Baseline eGFR from `pre_CRE_180days` |
| `eGFR_CRE_baseline` | Same baseline eGFR used in downstream analysis |
| `eGFR_CRE_1year` … `eGFR_CRE_10year` | eGFR at each yearly window (median pre-CRE in that year) |
| `eGFR_CRE_baseline_decrease_10` | `-eGFR_CRE_baseline / 10` (per 10-unit decrease) |
| `eGFR_CRE_1year_decrease_10` … `10year_decrease_10` | Per 10-unit decrease at each year |
| `ckd_stage_baseline`, `ckd_stage_1year` … `ckd_stage_10year` | Numeric CKD stage 1–5 from eGFR |

## 1.4 Derived — RPDR Pre-baseline Labs (within 180 days before ICI start)

| In Code | RPDR Group_Id | Units |
|---|---|---|
| `pre_CRE_180days` | `CRE` | mg/dL |
| `pre_HGB_180days` | `HGB` | g/dL |
| `pre_ALB_180days` | `ALB` | g/dL |
| `pre_mdeian_CRE_1year` … `pre_mdeian_CRE_10year` | `CRE` | Median CRE in each yearly window after ICI |

## 1.5 Derived — Comorbidities & Medications (on/before ICI)

| In Code | Source |
|---|---|
| `dm` | ICD-9/10 diabetes codes (`250.xx`, `E10.*`, `E11.*`, `E13.9`) |
| `htn` | ICD codes ± HTN medications, excluding pulmonary/venous HTN labels |
| `cad` | Cleaned CAD diagnosis name list |
| `cirrhosis` | Substring match on `Diagnosis_Name` |
| `esrd_kt` | ESRD / kidney transplant / dialysis ICD codes (used for exclusion) |
| `esrd_kt_after_ici` | Same codes but dates **after** `ICI_Dose_Date` (→ outcome) |
| `smoking` | Diagnosis names matching smoking/tobacco/cigarette/nicotine |
| `ppi` | Medication list (omeprazole, pantoprazole, esomeprazole, …) |
| `ace_arb` | ACEi/ARB medication list |
| `diu` | Loop / thiazide / K-sparing diuretic medication list |
| `steroids` | Systemic steroid medications, excluding inhaler/ointment/ophthalmic |
| `statins` | (Pre-computed in upstream master) |
| `allopurinol`, `sglt2i`, `glp1` | Medication lists (auxiliary) |

## 1.6 Derived — ICI Class & Co-administered Chemo (±30 days of ICI)

| In Code | Description |
|---|---|
| `ICI_Class` | PD1 / PDL1 / CTLA4 / Combination (from `ICI_Name`) |
| `Bevacizumab` | 1 if Bevacizumab/Avastin within ±30d of ICI |
| `Cisplatin`, `Carboplatin`, `Pemetrexed`, `Gemcitabine`, `Cetuximab`, `Trastuzumab` | Same pattern |
| `VEGF_TKI` | Sunitinib / Sorafenib / Pazopanib / Axitinib / Lenvatinib / Cabozantinib / Regorafenib / Tivozanib |
| `nephrotoxic_chemo` | 1 if any of Bevacizumab, Cisplatin, Carboplatin, Pemetrexed, Gemcitabine, Cetuximab, Trastuzumab == 1 |

## 1.7 Derived — Cancer Groups

| In Code | Mapping |
|---|---|
| `cancer_group` | gi / cutaneous-melanoma / gu / gyn / thoracic / other (from `cancer_type`) |

## 1.8 Derived — Outcomes (CKD / ESKD / Death)

All outcomes require a **90-day persistence rule** (the condition must hold
continuously ≥ 90 days) unless noted.

| In Code | Definition |
|---|---|
| `ckd_incidence` | First sustained (≥90d) episode with `pct_drop > 0.30` & `eGFR_CRE < 60` & `eGFR_CRE_baseline ≥ 60` |
| `ckd_incidence_date` | Start date of the first qualifying episode |
| `ckd_progression` | First sustained (≥90d) episode with `eGFR_CRE < 60` & `eGFR_CRE_baseline < 60` |
| `eskd_defination_1` | First sustained (≥90d) episode with `eGFR_CRE < 10` |
| `esrd_kt_after_ici` | 1 if ESRD/KT diagnosis date after `ICI_Dose_Date`; `esrd_kt_after_ici_date` = first post-ICI date |
| `eskd_composite` | `esrd_kt_after_ici == 1 \| eskd_defination_1 == 1` |
| `eskd_composite_date` | `pmin(esrd_kt_after_ici_date, eskd_defination_1_date)` |
| `ckd_composite` | `eskd_composite == 1 \| ckd_incidence == 1 \| ckd_progression == 1` |

## 1.9 Derived — Time Variables

| In Code | Description |
|---|---|
| `last_cre_date` | Most recent CRE lab date |
| `last_cre_days` | Days from ICI to `last_cre_date` |
| `death_days` / `time_to_death` | Days from ICI to `Date_Of_Death` |
| `time_to_death_status` | 0 = alive/censored, 2 = death |
| `time_to_eskd_composite` | Days from ICI to `eskd_composite_date` |
| `time_to_ckd_incidence` | Days from ICI to `ckd_incidence_date` |
| `time_to_ckd_progression` | Days from ICI to `ckd_progression_date` |
| `time_to_ckd_composite` | `pmin(time_to_eskd_composite, time_to_ckd_incidence, time_to_ckd_progression)` |

## 1.10 Derived — Competing Risk (CKD Composite vs Death)

| In Code | Description |
|---|---|
| `status_ckd_cmp` | Factor: `Censored` (0), `CKD Composite` (1), `Death` (2) |
| `time_to_ckd_cmp` | Event time matching `status_ckd_cmp` (CKD date, death date, or last-CRE date) |
| `time_to_ckd_cmp_year` | `time_to_ckd_cmp / 365` |
| `status_ckd_csh` | Cause-specific coding: 1 if `status_ckd_cmp == "CKD Composite"`, else 0 |

## 1.11 Derived — Follow-up & Year-Level Flags

| In Code | Description |
|---|---|
| `followed_1year` … `followed_10year` | 1 if `last_cre_days > 365 × N` |
| `ckd_incidence_1year` … `_10year` | 1 if `time_to_ckd_incidence < 365 × N` |
| `ckd_progression_1year` … `_10year` | Same pattern |
| `eskd_composite_1year` … `_10year` | Same pattern |
| `ckd_composite_1year` … `_10year` | Same pattern |

## 1.12 Derived — AKI

| In Code | Definition |
|---|---|
| `aki` | 1 if post-ICI CRE exceeds **1.5× median-CRE-in-prior-90-days** OR exceeds **lowest-CRE-in-prior-2-days** by ≥ 0.3 mg/dL |
| `aki_date` | First post-ICI date meeting the AKI condition |
| `aki_days` | Days from ICI to `aki_date` |
| `aki_before_ckd` | 1 if `aki_days <= time_to_ckd_composite` |

---

# 2 Setup & Libraries

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| label: libraries
rm(list = ls())
library(readxl)
library(writexl)
library(dplyr)
library(tidyverse)
library(tableone)
library(gtsummary)
library(survival)
library(ggplot2)
library(ggpubr)
library(splines)
library(lattice)
library(mice)
library(DescTools)
library(survminer)
library(finalfit)
library(tidycmprsk)
library(ggsurvfit)
library(sqldf)
library(table1)
```

---

# 3 Functions

## 3.1 standardize_colnames

```{r}
#| label: fn-standardize-colnames
standardize_colnames <- function(df) {
  names(df) <- gsub("[^[:alnum:]_]", "_", names(df))
  names(df) <- gsub("_+", "_", names(df))
  names(df) <- gsub("_$", "", names(df))
  return(df)
}
```

## 3.2 ckd_epi_gfr (race-free 2021 CKD-EPI)

```{r}
#| label: fn-ckd-epi-gfr
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  gfr <- 142 * min(c(creat/k, 1))^a *
               max(c(creat/k, 1))^-1.2 *
               0.9938^age * m
  return(gfr)
}
gfr_calc <- Vectorize(ckd_epi_gfr)
```

## 3.3 get_ckd_stage

```{r}
#| label: fn-get-ckd-stage
get_ckd_stage <- function(egfr) {
  if (is.na(egfr))      return(NA)
  else if (egfr >= 90)  return(1)
  else if (egfr >= 60)  return(2)
  else if (egfr >= 30)  return(3)
  else if (egfr >= 15)  return(4)
  else                  return(5)
}
```

## 3.4 pre_lab — closest lab on or before index date

```{r}
#| label: fn-pre-lab
pre_lab <- function(data, month, time, index_date, Group_Id) {
  index_date <- enquo(index_date)
  time_value <- if (month) time * 30 else time

  lab_name      <- paste0("pre_", Group_Id, "_", time_value, "days")
  lab_date_name <- paste0("pre_", Group_Id, "_", time_value, "days_date")

  data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date) - time_value,
           Seq_Date_Time <=  !!index_date) %>%
    arrange(EMPI, desc(Seq_Date_Time)) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    rename(!!lab_name      := Result,
           !!lab_date_name := Seq_Date_Time) %>%
    select(EMPI, !!lab_name, !!lab_date_name)
}
```

## 3.5 med_list — patients on a medication on/before index date

> **FIXED 2026-04-17 — Temporal leakage bug (flagged by Codex adversarial review).**
> Previous version only applied `Medication_Date <= index_date` when
> `excluded_names` was supplied; the `else` branch dropped the date guard, so
> exposures like `ppi`, `ace_arb`, and `diu` could include medications started
> AFTER ICI. This contaminated "baseline" covariates with post-ICI treatment
> history and biased downstream models. Date filter is now applied in **both**
> branches. Any `final_master_*.xlsx` built from the old version must be
> regenerated before use.

```{r}
#| label: fn-med-list
med_list <- function(names, index_date, excluded_names = NULL, data) {
  # Always enforce the pre-index date guard — do not remove.
  filtered <- data %>%
    filter(Medication_Date <= {{index_date}})

  if (!is.null(excluded_names)) {
    filtered %>%
      filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE) &
            !grepl(paste(excluded_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    filtered %>%
      filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  }
}
```

## 3.6 dia_list_icd / dia_list_icd_after_index / dia_list_name

```{r}
#| label: fn-dia-list-icd
dia_list_icd <- function(data, index_date, icd_code,
                         excluded_names = NULL, med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(icd_code,      collapse = "|"), Code,            ignore.case = TRUE),
            !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name,  ignore.case = TRUE)) %>%
      distinct(EMPI)
  }

  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }

  rbind(dia_result, med_result) %>% distinct(EMPI)
}
```

```{r}
#| label: fn-dia-list-icd-after-index
dia_list_icd_after_index <- function(data, index_date, icd_code,
                                     excluded_names = NULL,
                                     med_data = NULL, med_names = NULL) {
  dia_subset <- data %>%
    filter(Date > {{index_date}}) %>%
    filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE))

  if (!is.null(excluded_names)) {
    dia_subset <- dia_subset %>%
      filter(!grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE))
  }

  dia_result <- dia_subset %>%
    group_by(EMPI) %>%
    summarise(event_date = min(Date), .groups = "drop")

  med_result <- NULL
  if (!is.null(med_data) && !is.null(med_names)) {
    med_result <- med_data %>%
      filter(Medication_Date > {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      group_by(EMPI) %>%
      summarise(event_date = min(Medication_Date), .groups = "drop")
  }

  bind_rows(dia_result, med_result) %>%
    group_by(EMPI) %>%
    summarise(first_event_date = min(event_date), .groups = "drop")
}
```

```{r}
#| label: fn-dia-list-name
dia_list_name <- function(data, listed_names, index_date,
                          excluded_names = NULL, med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter( grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE),
             !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
      distinct(EMPI)
  }

  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }

  rbind(dia_result, med_result) %>% distinct(EMPI)
}
```

## 3.7 make_ckd_outcome_date — sustained-episode outcomes

Identifies the first episode where a condition (e.g., `eGFR < 60`) is sustained
for ≥ `days_threshold` days and returns both the binary flag and first-event date.

```{r}
#| label: fn-make-ckd-outcome-date
make_ckd_outcome_date <- function(data, id_var, date_var, condition_expr,
                                  days_threshold = 90, outcome_name = "default") {
  id_var      <- enquo(id_var)
  date_var    <- enquo(date_var)
  cond_q      <- enquo(condition_expr)
  outcome_sym <- ensym(outcome_name)
  outcome_date_name <- paste0(outcome_name, "_date")

  working <- data %>%
    arrange(!!id_var, !!date_var) %>%
    group_by(!!id_var) %>%
    mutate(
      .cond   = if_else(!!cond_q, 1L, 0L),
      .change = .cond != lag(.cond, default = first(.cond)),
      .run    = cumsum(.change)
    )

  episodes <- working %>%
    group_by(!!id_var, .run) %>%
    summarise(.code = first(.cond),
              .start = first(!!date_var),
              .last  = last(!!date_var),
              .groups = "drop") %>%
    group_by(!!id_var) %>%
    mutate(.end  = lead(.start),
           .end  = if_else(is.na(.end), .last, .end),
           .days = as.integer(.end - .start)) %>%
    ungroup()

  episodes %>%
    group_by(!!id_var) %>%
    summarise(
      has_event  = any(.code == 1L & .days >= days_threshold),
      event_date = if (any(.code == 1L & .days >= days_threshold))
                      min(.start[.code == 1L & .days >= days_threshold])
                   else as.Date(NA),
      .groups = "drop"
    ) %>%
    mutate(!!outcome_sym := as.integer(has_event)) %>%
    rename(!!outcome_date_name := event_date) %>%
    select(-has_event)
}
```

## 3.8 create_aki — KDIGO-style AKI from serial CRE

```{r}
#| label: fn-create-aki
create_aki <- function(data,
                       index_date = ICI_Dose_Date,
                       Group_Id   = EMPI,
                       median_window_days = 90,
                       lowest_window_days = 2) {
  index_date <- enquo(index_date)
  Group_Id   <- enquo(Group_Id)

  df <- data %>%
    filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    rename(cre_date = Seq_Date_Time, cre = Result) %>%
    select(EMPI, cre, cre_date, !!index_date) %>%
    arrange(!!Group_Id, cre_date)

  df %>%
    filter(cre_date > !!index_date) %>%
    group_by(!!Group_Id) %>%
    mutate(
      median_cre_90days = sapply(seq_along(cre), function(i) {
        vals <- cre[cre_date >= cre_date[i] - median_window_days &
                    cre_date <= cre_date[i] - 1]
        if (length(vals) == 0) NA_real_ else median(vals, na.rm = TRUE)
      }),
      lowest_cre_2days = sapply(seq_along(cre), function(i) {
        vals <- cre[cre_date >= cre_date[i] - lowest_window_days &
                    cre_date <= cre_date[i] - 1]
        if (length(vals) == 0) NA_real_ else min(vals, na.rm = TRUE)
      }),
      aki_flag = (!is.na(median_cre_90days) & cre > 1.5 * median_cre_90days) |
                 (!is.na(lowest_cre_2days)  & cre - lowest_cre_2days >= 0.3)
    ) %>%
    filter(aki_flag) %>%
    arrange(!!Group_Id, cre_date) %>%
    summarise(aki      = 1L,
              aki_date = min(cre_date), .groups = "drop")
}
```

## 3.9 left_join_all — fold a list of dfs into master by key

```{r}
#| label: fn-left-join-all
left_join_all <- function(master, df_list, by) {
  for (df in df_list) master <- merge(master, df, by = by, all.x = TRUE)
  master
}
```

---

# 4 Data Loading

## 4.1 ICI starts master

```{r}
#| label: load-ici-master
master <- read_excel('/Users/to909/Desktop/Meg/Kavita ICI CKD/ICI_starts_downloaded_1127.xlsx')
master <- standardize_colnames(master)
master <- master %>% distinct(EMPI_MRN, .keep_all = TRUE) %>% rename(EMPI = EMPI_MRN)
master$ICI_Dose_Date <- as.Date(master$ICI_Dose_Date)
master_list <- master %>% select(EMPI, ICI_Dose_Date)
```

## 4.2 Load RPDR files (Lab / Dem / Med / Dia)

```{r}
#| label: load-rpdr
rpdr_path <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR"

lab <- read.table(file.path(rpdr_path, "mes98_093025104659807332_Lab.txt"),
                  sep = "|", fill = TRUE, quote = "", header = TRUE) %>%
  select(EMPI, EPIC_PMRN, Seq_Date_Time, Result, Group_Id, Test_Description)
lab$Seq_Date_Time <- as.Date(lab$Seq_Date_Time, "%m/%d/%Y")

dem <- read.table(file.path(rpdr_path, "mes98_093025104659807332_Dem.txt"),
                  sep = "|", fill = TRUE, quote = "", header = TRUE)
dem$Date_of_Birth <- as.Date(dem$Date_of_Birth, "%m/%d/%Y")
dem$Date_Of_Death <- as.Date(dem$Date_Of_Death, "%m/%d/%Y")

med <- read.table(file.path(rpdr_path, "mes98_093025104659807332_Med.txt"),
                  sep = "|", fill = TRUE, quote = "", header = TRUE) %>%
  select(EMPI, EPIC_PMRN, Medication_Date, Medication)
med$Medication_Date <- as.Date(med$Medication_Date, "%m/%d/%Y")

dia <- read.table(file.path(rpdr_path, "mes98_093025104659807332_Dia.txt"),
                  sep = "|", fill = TRUE, quote = "", header = TRUE) %>%
  select(EMPI, EPIC_PMRN, Code, Date, Diagnosis_Name)
dia$Date <- as.Date(dia$Date, "%m/%d/%Y")
```

## 4.3 Merge RPDR with master index date

```{r}
#| label: load-merge-index
master_lab <- lab %>% merge(master_list, by = "EMPI", all.x = TRUE)
master_med <- med %>% merge(master_list, by = "EMPI", all.x = TRUE)
master_dia <- dia %>% merge(master_list, by = "EMPI", all.x = TRUE)
```

---

# 5 Data Cleaning & Variable Creation

## 5.1 Demographics & Race / Ethnic_Group

```{r}
#| label: clean-demographics
dems <- dem %>%
  mutate(
    Race1 = ifelse(is.na(Race1), "Other/Unknown", Race1),
    Race  = case_when(
      Race1 == "White" ~ "White",
      Race1 == "Black" ~ "Black",
      Race1 == "Asian" ~ "Asian",
      TRUE             ~ "Other/Unknown"
    ),
    Ethnic_Group = case_when(
      Ethnic_Group == "HISPANIC"     ~ "Hispanic",
      Ethnic_Group == "Non Hispanic" ~ "Non_hispanic",
      TRUE                           ~ "Other"
    ),
    Ethnic_Group = case_when(
      Race != "Black" & Language == "Spanish" ~ "Hispanic",
      TRUE                                     ~ Ethnic_Group
    )
  ) %>%
  distinct(EMPI, .keep_all = TRUE)

master <- master %>%
  merge(dems, by = "EMPI") %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  mutate(male = if_else(Gender_Legal_Sex == "Male", 1, 0))
```

## 5.2 Last creatinine date & follow-up days

```{r}
#| label: clean-last-cre
last_cre <- master_lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(Result = as.numeric(Result)) %>%
  arrange(EMPI, desc(Seq_Date_Time)) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  rename(last_cre_date = Seq_Date_Time) %>%
  select(EMPI, last_cre_date)

master <- master %>% merge(last_cre, by = "EMPI", all.x = TRUE) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  mutate(
    last_date = as.Date(ifelse(is.na(Date_Of_Death), last_cre_date, Date_Of_Death)),
    last_days = as.numeric(last_date - ICI_Dose_Date)
  )
```

## 5.3 Medication & diagnosis patient lists (on/before ICI)

```{r}
#| label: clean-med-lists
ppi_names     <- c("omeprazole","Prilosec","Yosprala","lansoprazole","Prevacid",
                   "dexlansoprazole","Dexilent","rabeprazole","Aciphex","pantoprazole",
                   "Protonix","esomeprazole","Nexium","Vimovo","Zegerid","dexilant")
ace_arb_names <- c("benazepril","lotensin","captopril","capoten","enalapril","vasotech",
                   "epaned","lexxel","fosinopril","monopril","lisinopril","prinvil",
                   "zestril","qbrelis","moexipril","univasc","perindopril","aceon",
                   "quinapril","accupril","ramipril","altace","trandolapril","mavik",
                   "azilsartan","edarbi","candesartan","atacand","eprosartan","teveten",
                   "irbesartan","avapro","telmisartan","micardis","valsartan","diovan",
                   "losartan","cozaar","olmesartan","benicar","lotrel","hyzaar")
diuretics_names <- c("bumetanide","bumex","ethacrynic acid","edecrin","furosemide","Lasix",
                     "torsemide","demadex","hydrochlorothiazide","hctz","chlorathiazide",
                     "chlorthalidone","diuril","indapamide","metolazone","aldactone",
                     "amiloride","dyazide","eplerenone","maxzide","spironolactone","triamterene")
steroids_names   <- c("prednisone","methylpred","Deltasone","Orasone","budesonide","Entocort",
                      "methylprednisolone","prednisolone","Millipred","dexamethasone",
                      "Ozurdex","Maxidex","DexPak")
no_steroid_names <- c("Inhaler","Nebulization","Ointment","Ophthalmic","Symbicort")

ppi     <- med_list(data = master_med, names = ppi_names,        index_date = ICI_Dose_Date)
ace_arb <- med_list(data = master_med, names = ace_arb_names,    index_date = ICI_Dose_Date)
diu     <- med_list(data = master_med, names = diuretics_names,  index_date = ICI_Dose_Date)
steroids <- med_list(data = master_med, names = steroids_names,
                     excluded_names = no_steroid_names,           index_date = ICI_Dose_Date)
```

## 5.4 Diagnosis patient lists (on/before ICI)

```{r}
#| label: clean-dia-lists
esrd_kt_codes <- c("585.6","N18.6","V45.1","Z94.0","Z99.2","996.81","T86.1","T86.10",
                   "T86.11","T86.12","T86.13","T86.19","N18.0","V56.1","Z49.0","Z49.1","Z49.2")
diabetes_codes <- c("250.00","250.01","250.02","250.03","250.10","250.11","250.12",
                    "250.13","250.40","250.41","250.42","250.43",
                    "E10.9","E11.9","E10.65","E11.65","E10.21","E11.21","E10.22","E11.22",
                    "E10.29","E11.29","E10.40","E11.40","E10.49","E11.49","E13.9")
htn_codes     <- c("401.1","401.0","401.9","I10","I15.1","405.99","I15.9","I15.8",
                   "405.09","997.91","405.91","I15.0","I15.2","405.01")
no_htn_list   <- c("Pulmonary Embolism","Borderline hypertension","Labile hypertension",
                   "venous hypertension")

esrd_kt           <- dia_list_icd(master_dia, icd_code = esrd_kt_codes,
                                   index_date = ICI_Dose_Date)
esrd_kt_after_ici <- dia_list_icd_after_index(master_dia, icd_code = esrd_kt_codes,
                                               index_date = ICI_Dose_Date) %>%
                     rename(esrd_kt_after_ici_date = first_event_date)

dm  <- dia_list_icd(master_dia, icd_code = diabetes_codes, index_date = ICI_Dose_Date)
htn <- dia_list_icd(master_dia, icd_code = htn_codes,
                    excluded_names = no_htn_list,
                    index_date = ICI_Dose_Date)

cirrhosis <- master_dia %>%
  filter(Date <= ICI_Dose_Date) %>%
  filter(grepl("cirrhosis", Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)

cad <- dia_list_name(master_dia,
                     listed_names = c("coronary artery disease","coronary atherosclerosis",
                                      "angina","ischemic heart disease","atherosclerosis"),
                     index_date = ICI_Dose_Date)

smoking <- master_dia %>%
  filter(grepl("smoking|tobacco|cigarette|cigarettes|nicotine|smoke",
               Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)
```

## 5.5 Fold comorbidities into master

```{r}
#| label: clean-fold-comorbidities
all_dem <- sqldf("
  select distinct master.EMPI,
         case when diu.EMPI      is not null then 1 else 0 end as diu,
         case when ace_arb.EMPI  is not null then 1 else 0 end as ace_arb,
         case when esrd_kt.EMPI  is not null then 1 else 0 end as esrd_kt,
         case when dm.EMPI       is not null then 1 else 0 end as dm,
         case when htn.EMPI      is not null then 1 else 0 end as htn,
         case when ppi.EMPI      is not null then 1 else 0 end as ppi,
         case when steroids.EMPI is not null then 1 else 0 end as steroids,
         case when smoking.EMPI  is not null then 1 else 0 end as smoking,
         case when cad.EMPI      is not null then 1 else 0 end as cad
  from master
    left join diu      on diu.EMPI      = master.EMPI
    left join dm       on dm.EMPI       = master.EMPI
    left join ace_arb  on ace_arb.EMPI  = master.EMPI
    left join esrd_kt  on esrd_kt.EMPI  = master.EMPI
    left join htn      on htn.EMPI      = master.EMPI
    left join steroids on steroids.EMPI = master.EMPI
    left join cad      on cad.EMPI      = master.EMPI
    left join smoking  on smoking.EMPI  = master.EMPI
    left join ppi      on ppi.EMPI      = master.EMPI
")
```

## 5.6 Pre-baseline labs & baseline eGFR

```{r}
#| label: clean-pre-labs
pre_CRE_180days <- pre_lab(master_lab, month = FALSE, time = 180,
                            index_date = ICI_Dose_Date, Group_Id = "CRE")
pre_HGB_180days <- pre_lab(master_lab, month = FALSE, time = 180,
                            index_date = ICI_Dose_Date, Group_Id = "HGB")
pre_ALB_180days <- pre_lab(master_lab, month = FALSE, time = 180,
                            index_date = ICI_Dose_Date, Group_Id = "ALB")

dem_lab_list <- list(all_dem, pre_CRE_180days, pre_HGB_180days, pre_ALB_180days)
all_outcome_lab_dem <- left_join_all(master, dem_lab_list, by = "EMPI")

all_outcome_lab_dem <- all_outcome_lab_dem %>%
  mutate(age = round(as.numeric(difftime(ICI_Dose_Date, Date_of_Birth,
                                         units = "days")) / 365))

all_outcome_lab_dem$eGFR_cre <- mapply(ckd_epi_gfr,
                                       all_outcome_lab_dem$pre_CRE_180days,
                                       all_outcome_lab_dem$male,
                                       all_outcome_lab_dem$age)
```

## 5.7 Longitudinal CRE after baseline & CKD outcomes

```{r}
#| label: clean-longitudinal-cre
master_lab_with_cre <- master_lab %>%
  merge(all_outcome_lab_dem %>%
          select(EMPI, male, age, pre_CRE_180days, pre_CRE_180days_date),
        by = "EMPI", all.x = TRUE)

master_lab_cre_after_baseline <- master_lab_with_cre %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > pre_CRE_180days_date)

master_lab_cre_after_baseline$eGFR_CRE_baseline <-
  mapply(gfr_calc,
         master_lab_cre_after_baseline$pre_CRE_180days,
         master_lab_cre_after_baseline$male,
         master_lab_cre_after_baseline$age)

master_lab_cre_after_baseline <- master_lab_cre_after_baseline %>%
  rename(cre = Result, cre_date = Seq_Date_Time)

master_lab_cre_after_baseline$eGFR_CRE <-
  mapply(gfr_calc, master_lab_cre_after_baseline$cre,
         master_lab_cre_after_baseline$male,
         master_lab_cre_after_baseline$age)

master_lab_cre_after_baseline <- master_lab_cre_after_baseline %>%
  rename(eGFR_CRE_date         = cre_date,
         eGFR_CRE_baseline_date = pre_CRE_180days_date)

test <- master_lab_cre_after_baseline %>%
  select(EMPI, eGFR_CRE_baseline, eGFR_CRE_baseline_date,
         eGFR_CRE, eGFR_CRE_date) %>%
  mutate(pct_drop = round((eGFR_CRE_baseline - eGFR_CRE) / eGFR_CRE_baseline, 3)) %>%
  arrange(EMPI, eGFR_CRE_date)

ckd_incidence <- make_ckd_outcome_date(
  data = test, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = pct_drop > 0.30 & eGFR_CRE < 60 & eGFR_CRE_baseline >= 60,
  days_threshold = 90, outcome_name = "ckd_incidence")

eskd_defination_1 <- make_ckd_outcome_date(
  data = test, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = eGFR_CRE < 10,
  days_threshold = 90, outcome_name = "eskd_defination_1")

ckd_progression <- make_ckd_outcome_date(
  data = test, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = eGFR_CRE < 60 & eGFR_CRE_baseline < 60,
  days_threshold = 90, outcome_name = "ckd_progression")
```

## 5.8 Assemble final_master & cancer type merge

```{r}
#| label: clean-assemble-final-master
final_master <- all_outcome_lab_dem %>%
  merge(ckd_incidence,     by = "EMPI", all.x = TRUE) %>%
  merge(eskd_defination_1, by = "EMPI", all.x = TRUE) %>%
  merge(ckd_progression,   by = "EMPI", all.x = TRUE) %>%
  merge(esrd_kt_after_ici, by = "EMPI", all.x = TRUE) %>%
  mutate(esrd_kt_after_ici = ifelse(!is.na(esrd_kt_after_ici_date), 1, 0))

cancer_type <- readxl::read_xlsx(
  '/Users/to909/Desktop/Meg/Kavita ICI CKD/final_master_1202.xlsx') %>%
  select(EMPI, Tumor_Category)

final_master <- final_master %>%
  select(-Tumor_Category) %>%
  merge(cancer_type, by = "EMPI", all.x = TRUE) %>%
  rename(cancer_type = Tumor_Category)
```

## 5.9 Cohort exclusions

```{r}
#| label: clean-exclusions
cat("Before exclusions:", nrow(final_master), "\n")

final_master <- final_master %>%
  filter(ICI_Dose_Date < "2024-03-31") %>%       # must have ≥ follow-up window
  filter(!is.na(pre_CRE_180days)) %>%            # must have baseline creatinine
  filter(esrd_kt == 0) %>%                       # exclude pre-existing ESRD/KT
  filter(eGFR_cre >= 10) %>%                     # exclude pre-existing stage-5
  filter(is.na(Date_Of_Death) | Date_Of_Death >= ICI_Dose_Date)  # RPDR date errors

cat("After exclusions (target ~17,718):", nrow(final_master), "\n")
```

## 5.10 Co-administered chemo within ±30 days of ICI

```{r}
#| label: clean-chemo-indicators
master_med_window <- med %>%
  merge(final_master %>% select(EMPI, ICI_Dose_Date), by = "EMPI")

med_list_window <- function(names, index_date, data) {
  data %>%
    filter(Medication_Date >= {{index_date}} - 30,
           Medication_Date <= {{index_date}} + 30) %>%
    filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
    distinct(EMPI)
}

Bevacizumab  <- med_list_window(c("Bevacizumab","Avastin"),           ICI_Dose_Date, master_med_window)
Cisplatin    <- med_list_window(c("Cisplatin"),                        ICI_Dose_Date, master_med_window)
Carboplatin  <- med_list_window(c("Carboplatin"),                      ICI_Dose_Date, master_med_window)
Pemetrexed   <- med_list_window(c("Pemetrexed","Alimta"),              ICI_Dose_Date, master_med_window)
Gemcitabine  <- med_list_window(c("Gemcitabine","Gemzar"),             ICI_Dose_Date, master_med_window)
Cetuximab    <- med_list_window(c("Cetuximab","Erbitux"),              ICI_Dose_Date, master_med_window)
Trastuzumab  <- med_list_window(c("Trastuzumab","Herceptin"),          ICI_Dose_Date, master_med_window)
VEGF_TKI     <- med_list_window(c("Sunitinib","Sutent","Sorafenib","Nexavar","Pazopanib",
                                   "Votrient","Axitinib","Inlyta","Lenvatinib","Lenvima",
                                   "Cabozantinib","Cabometyx","Cometriq","Regorafenib",
                                   "Stivarga","Tivozanib","Fotivda"),
                                 ICI_Dose_Date, master_med_window)

final_master <- final_master %>% rename(age_ici = age)

final_master <- sqldf("
  select distinct final_master.*,
         case when Bevacizumab.EMPI is not null then 1 else 0 end as Bevacizumab,
         case when Cisplatin.EMPI   is not null then 1 else 0 end as Cisplatin,
         case when Carboplatin.EMPI is not null then 1 else 0 end as Carboplatin,
         case when Pemetrexed.EMPI  is not null then 1 else 0 end as Pemetrexed,
         case when Gemcitabine.EMPI is not null then 1 else 0 end as Gemcitabine,
         case when Cetuximab.EMPI   is not null then 1 else 0 end as Cetuximab,
         case when Trastuzumab.EMPI is not null then 1 else 0 end as Trastuzumab,
         case when VEGF_TKI.EMPI    is not null then 1 else 0 end as VEGF_TKI
  from final_master
    left join Bevacizumab on Bevacizumab.EMPI = final_master.EMPI
    left join Cisplatin   on Cisplatin.EMPI   = final_master.EMPI
    left join Carboplatin on Carboplatin.EMPI = final_master.EMPI
    left join Pemetrexed  on Pemetrexed.EMPI  = final_master.EMPI
    left join Gemcitabine on Gemcitabine.EMPI = final_master.EMPI
    left join Cetuximab   on Cetuximab.EMPI   = final_master.EMPI
    left join Trastuzumab on Trastuzumab.EMPI = final_master.EMPI
    left join VEGF_TKI    on VEGF_TKI.EMPI    = final_master.EMPI
")

final_master <- final_master %>%
  mutate(nephrotoxic_chemo = if_else(
    if_any(c(Bevacizumab, Cisplatin, Carboplatin, Pemetrexed,
             Gemcitabine, Cetuximab, Trastuzumab), ~ . == 1), 1L, 0L))
```

## 5.11 AKI from serial creatinine

```{r}
#| label: clean-aki
aki_df <- create_aki(master_lab,
                     index_date = ICI_Dose_Date,
                     Group_Id   = EMPI,
                     median_window_days = 90,
                     lowest_window_days = 2)

final_master <- final_master %>%
  merge(aki_df, by = "EMPI", all.x = TRUE) %>%
  mutate(
    aki      = if_else(is.na(aki), 0, aki),
    aki_days = as.numeric(aki_date - ICI_Dose_Date)
  )
```

## 5.12 ICI class mapping

```{r}
#| label: clean-ici-class
final_master <- final_master %>%
  mutate(ICI_Class = case_when(
    ICI_Name %in% c("pembrolizumab","nivolumab","cemiplimab",
                    "Nivolumab","Dostarlimab","Cemiplimab") ~ "PD1",
    ICI_Name %in% c("atezolizumab","avelumab","durvalumab")  ~ "PDL1",
    ICI_Name == "ipilimumab"                                 ~ "CTLA4",
    ICI_Name == "ipilimumab AND nivolumab"                   ~ "Combination",
    TRUE                                                      ~ NA_character_
  ))
```

## 5.13 Cancer groups

```{r}
#| label: clean-cancer-group
final_master <- final_master %>%
  mutate(cancer_group = case_when(
    cancer_type == "gi"                    ~ "gi",
    cancer_type == "cutaneous / melanoma"  ~ "cutaneous / melanoma",
    cancer_type == "gu"                    ~ "gu",
    cancer_type == "gyn"                   ~ "gyn",
    cancer_type == "thoracic"              ~ "thoracic",
    TRUE                                    ~ "other"
  ))
```

## 5.14 Composite outcomes & time variables

```{r}
#| label: clean-composite-outcomes
final_master <- final_master %>%
  mutate(
    eskd_composite       = ifelse(esrd_kt_after_ici == 1 | eskd_defination_1 == 1, 1, 0),
    eskd_composite       = ifelse(is.na(eskd_composite), 0, eskd_composite),
    eskd_composite_date  = pmin(esrd_kt_after_ici_date, eskd_defination_1_date, na.rm = TRUE),
    time_to_eskd_composite  = as.numeric(eskd_composite_date  - ICI_Dose_Date),
    time_to_ckd_incidence   = as.numeric(ckd_incidence_date   - ICI_Dose_Date),
    time_to_ckd_progression = as.numeric(ckd_progression_date - ICI_Dose_Date),

    ckd_composite = ifelse(eskd_composite == 1 | ckd_incidence == 1 |
                           ckd_progression == 1, 1, 0),
    ckd_composite = ifelse(is.na(ckd_composite), 0, ckd_composite),
    time_to_ckd_composite = pmin(time_to_eskd_composite,
                                 time_to_ckd_incidence,
                                 time_to_ckd_progression, na.rm = TRUE),

    last_cre_days          = as.numeric(last_cre_date - ICI_Dose_Date),
    death_days             = as.numeric(Date_Of_Death - ICI_Dose_Date),
    time_to_death          = death_days,
    time_to_death_status   = ifelse(!is.na(death_days), 2, 0)
  )
```

## 5.15 Competing-risk status (CKD Composite vs Death)

```{r}
#| label: clean-competing-risk
final_master <- final_master %>%
  mutate(
    status_ckd_cmp = case_when(
      (ckd_composite == 1) & (time_to_death_status == 2) ~
        ifelse(time_to_ckd_composite < time_to_death, 1, 2),
      (ckd_composite == 1 & time_to_death_status == 0)   ~ 1,
      (ckd_composite == 0 & time_to_death_status == 2)   ~ 2,
      TRUE                                                ~ 0
    ),
    time_to_ckd_cmp = case_when(
      status_ckd_cmp == 1 ~ time_to_ckd_composite,
      status_ckd_cmp == 2 ~ time_to_death,
      status_ckd_cmp == 0 ~ last_cre_days
    ),
    status_ckd_cmp = factor(status_ckd_cmp, levels = c(0, 1, 2),
                            labels = c("Censored", "CKD Composite", "Death")),
    time_to_ckd_cmp_year = time_to_ckd_cmp / 365,
    status_ckd_csh       = ifelse(status_ckd_cmp == "CKD Composite", 1, 0)
  )

final_master <- final_master %>%
  mutate(aki_before_ckd = if_else(aki_days <= time_to_ckd_composite, 1, 0))
```

## 5.16 Yearly follow-up flags

```{r}
#| label: clean-followed-flags
final_master <- final_master %>%
  mutate(
    followed_1year  = ifelse(last_cre_days > 365 * 1, 1, 0),
    followed_2year  = ifelse(last_cre_days > 365 * 2, 1, 0),
    followed_3year  = ifelse(last_cre_days > 365 * 3, 1, 0),
    followed_4year  = ifelse(last_cre_days > 365 * 4, 1, 0),
    followed_5year  = ifelse(last_cre_days > 365 * 5, 1, 0),
    followed_6year  = ifelse(last_cre_days > 365 * 6, 1, 0),
    followed_7year  = ifelse(last_cre_days > 365 * 7, 1, 0),
    followed_8year  = ifelse(last_cre_days > 365 * 8, 1, 0),
    followed_9year  = ifelse(last_cre_days > 365 * 9, 1, 0),
    followed_10year = ifelse(last_cre_days > 365 * 10, 1, 0)
  )
```

## 5.17 Yearly median creatinine → yearly eGFR → yearly CKD stage

Compute the per-patient median CRE within each 365-day window after ICI,
then derive eGFR and CKD stage at each year.

```{r}
#| label: clean-yearly-median-cre
yearly_cre <- master_lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(
    Result         = as.numeric(Result),
    days_since_ici = as.numeric(Seq_Date_Time - ICI_Dose_Date),
    year_bucket    = ceiling(days_since_ici / 365)
  ) %>%
  filter(year_bucket >= 1, year_bucket <= 10) %>%
  group_by(EMPI, year_bucket) %>%
  summarise(median_cre = median(Result, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from  = year_bucket,
    values_from = median_cre,
    names_glue  = "pre_mdeian_CRE_{year_bucket}year"
  )

# Guarantee all 10 columns exist even if no patient has labs in that year
needed_yearly_cols <- paste0("pre_mdeian_CRE_", 1:10, "year")
for (col in setdiff(needed_yearly_cols, names(yearly_cre))) {
  yearly_cre[[col]] <- NA_real_
}

final_master <- final_master %>% merge(yearly_cre, by = "EMPI", all.x = TRUE)
```

```{r}
#| label: clean-yearly-egfr
for (y in 1:10) {
  col_cre  <- paste0("pre_mdeian_CRE_", y, "year")
  col_egfr <- paste0("eGFR_CRE_",       y, "year")
  final_master[[col_egfr]] <- mapply(gfr_calc,
                                     final_master[[col_cre]],
                                     final_master$male,
                                     final_master$age_ici)
}

# Fill eGFR at each year with baseline if missing
final_master <- final_master %>%
  mutate(across(any_of(paste0("eGFR_CRE_", 1:10, "year")),
                ~ coalesce(., eGFR_CRE_baseline)))

# Numeric CKD stage from eGFR
final_master$ckd_stage_baseline <- sapply(final_master$eGFR_CRE_baseline, get_ckd_stage)
for (y in 1:10) {
  col_egfr  <- paste0("eGFR_CRE_", y, "year")
  col_stage <- paste0("ckd_stage_", y, "year")
  final_master[[col_stage]] <- sapply(final_master[[col_egfr]], get_ckd_stage)
}
```

## 5.17 Per-10-unit decrease scaling & year-level outcome flags

```{r}
#| label: clean-year-outcome-flags
final_master <- final_master %>%
  mutate(
    year_10                        = age_ici / 10,
    eGFR_CRE_baseline_decrease_10  = (eGFR_CRE_baseline * -1) / 10
  )

for (y in 1:10) {
  col_egfr <- paste0("eGFR_CRE_",            y, "year")
  col_dec  <- paste0("eGFR_CRE_",            y, "year_decrease_10")
  final_master[[col_dec]] <- (final_master[[col_egfr]] * -1) / 10
}

cutoffs <- 1:10
for (y in cutoffs) {
  d <- 365 * y
  final_master[[paste0("eskd_composite_",    y, "year")]] <- ifelse(final_master$time_to_eskd_composite   < d, 1, 0)
  final_master[[paste0("ckd_incidence_",     y, "year")]] <- ifelse(final_master$time_to_ckd_incidence    < d, 1, 0)
  final_master[[paste0("ckd_progression_",   y, "year")]] <- ifelse(final_master$time_to_ckd_progression  < d, 1, 0)
  final_master[[paste0("ckd_composite_",     y, "year")]] <- ifelse(final_master$time_to_ckd_composite    < d, 1, 0)
}

final_master <- final_master %>%
  mutate(across(matches("(ckd_incidence|ckd_progression|eskd_composite|ckd_composite).*year$"),
                ~ replace_na(., 0))) %>%
  mutate(across(matches("(ckd_incidence|ckd_progression|eskd_composite).*year$"),
                ~ factor(replace_na(., 0))))
```

## 5.18 Factor conversion for categorical variables

```{r}
#| label: clean-factor-vars
num_var <- c("age_ici","pre_CRE_180days","year_10",
             "pre_HGB_180days","pre_ALB_180days","eGFR_CRE_baseline")
all_var <- c("age_ici","year_10","Gender_Legal_Sex","male","Race","Ethnic_Group",
             "ckd_incidence","ckd_progression","eskd_composite","ckd_composite",
             paste0("ckd_stage_", 1:10, "year"),
             "dm","htn","cad","cirrhosis","esrd_kt",
             "ace_arb","diu","ppi","steroids","smoking","statins",
             "ICI_Class","nephrotoxic_chemo",
             "pre_CRE_180days","pre_HGB_180days","pre_ALB_180days",
             "eGFR_CRE_baseline","ckd_stage_baseline",
             paste0("ckd_composite_", 1:5, "year"),
             "Bevacizumab","Cisplatin","Carboplatin","Pemetrexed",
             "Gemcitabine","Cetuximab","Trastuzumab",
             "cancer_type","cancer_group","VEGF_TKI","aki")

cat_var <- setdiff(all_var, num_var)
final_master <- final_master %>% mutate_at(intersect(cat_var, names(.)), factor)

if ("ICI_Class" %in% names(final_master)) {
  final_master$ICI_Class <- relevel(final_master$ICI_Class, ref = "PD1")
}
```

## 5.19 Per-year filtered cohorts (for year-specific analyses)

```{r}
#| label: clean-year-cohorts
final_master_1year <- final_master %>% filter(followed_1year == 1)
final_master_2year <- final_master %>% filter(followed_2year == 1)
final_master_3year <- final_master %>% filter(followed_3year == 1)
final_master_4year <- final_master %>% filter(followed_4year == 1)
final_master_5year <- final_master %>% filter(followed_5year == 1)
```

---

# 6 Save cleaned master

```{r}
#| label: save-final-master
saveRDS(final_master, "final_master.rds")
```

---

# 7 Cohort Exclusion Ledger (documentation)

| Step | N |
|---|---|
| Raw Excel (`ICI_starts_downloaded_1127.xlsx`) | 25,397 |
| After merging with demographics | 20,383 |
| After removing no baseline CRE | 18,319 |
| After `esrd_kt == 0` and `eGFR_CRE_baseline >= 10` | 18,157 |
| After `last_cre_date >= ICI_Dose_Date` (labs after ICI) | 17,937 |
| After removing 216 placebo | 17,721 |
| After removing 3 death-before-ICI RPDR errors | 17,718 |
